---
title: "SWB Metrics Report"
subtitle: "AIM-AHEAD non-FISMA"
output: pdf_document
---



```{r include = FALSE}
#Download all entries from the following DynamoDB tables: 
#- aim-prod-va-sw-Users
#- aim-prod-va-sw-EnvironmentsSc
#- Studies
#- StudyPermissions

users <- read.csv('aa-nonfisma-users.csv')
workspaces <- read.csv('aa-nonfisma-environments.csv')
studies <- read.csv('aa-nonfisma-studies.csv')
studyperm <- read.csv('aa-nonfisma-studypermissions.csv')

library(tidyverse)
library(knitr)

clean_list <- function(vec, sep, rep, newsep){
  vec = gsub(pattern = sep, replacement = rep, x = vec)
  vec = gsub(pattern = '[[:punct:]]', replacement = '', x = vec)
  vec = gsub(pattern = paste0('S', rep), replacement = newsep, x = vec)
  vec = gsub(pattern = paste0('^', newsep), replacement = '', x = vec)
  return(vec)
}

# keep testing to show behind the scenes work
remove_testing = FALSE

```

```{r include = FALSE}
# project groupings are manually generated

project_groupings <- c('testingresearch' = 'Testing',
          'DSTCPRIME' = 'AIM-AHEAD Fellowships',
          'AISC' = 'Teaching',
          'ResearchFellowship' = 'AIM-AHEAD Fellowships',
          'testingteaching' = 'Testing',
          'BMI705' = 'Teaching', 
          'testBMI703' = 'Testing', 
          'tianxiharmonization' = 'Other Projects',
          'tianxiharmonization2' = 'Other Projects',
          'BMI2023' = 'Teaching', 
          'AVLdev' = 'Testing', 
          'PilotCartwrightGunviolence' = 'AIM-AHEAD Pilot Projects',
          'PilotTamilMedCognetics' = 'AIM-AHEAD Pilot Projects',
          'LeadershipFellowship' = 'AIM-AHEAD Fellowships',
          'NorthStarCallForProposal' = 'AIM-AHEAD North Star Call for Proposals') %>% 
  as.data.frame() 
colnames(project_groupings) <- "Project_Group"
project_groupings$projectId = rownames(project_groupings)

```

```{r include = FALSE}

users_clean <- users %>% 
  select(uid, firstName, lastName, email, username, applyReason, 
         projectId, status,  userRole,
         createdAt, acceptedTerms, identityProviderName, isAdmin) %>%
  mutate(createdAt = date(createdAt), 
         acceptedTerms = ifelse(acceptedTerms == '', NA, acceptedTerms),
         acceptedTerms = date(acceptedTerms),
         projectId = clean_list(projectId, sep = ':', rep = 'XXX', newsep = ', ')) 

user_mapping <- users_clean %>% 
  select(uid, username) %>%
  mutate(uid = clean_list(uid, '', '', ''))

```

```{r include = FALSE}


# workspace mapping manually generated based on default configurations
workspace_mapping <- read.csv('workspace_mapping.csv') %>% 
  mutate(workspace_name = tolower(workspace_name),
         workspace_name = clean_list(workspace_name, sep = '', rep = '', newsep = ''))

workspaces_clean <- workspaces %>% 
  select(id, name, createdAt, createdBy, description, outputs,
         envTypeConfigId, indexId,  projectId, status, studyIds, updatedAt) %>%
  mutate(createdAt = date(createdAt),
         updatedAt = date(updatedAt), 
         projectId = gsub(pattern = '[[:punct:]]', replacement = '', x = projectId),
         studyIds = clean_list(studyIds, sep = ':', rep = 'XXX', newsep = ', '),
         envTypeConfigId = tolower(envTypeConfigId), 
         envTypeConfigId = clean_list(envTypeConfigId, sep = '', rep = '', newsep = '')) %>%
  left_join(users_clean %>% select(uid, username), by = c('createdBy' = 'uid')) %>%
  left_join(workspace_mapping, by = c('envTypeConfigId' = 'workspace_name'))   %>% 
  left_join(project_groupings) 

if(remove_testing){
  workspaces_clean <- workspaces_clean %>% filter(Project_Group != "Testing")
}


```

```{r include = FALSE}
studies_clean <- studies %>%
  select(id, accessType, bucket, category, createdAt, description, folder, name, projectId, updatedAt) %>%
  mutate(createdAt = date(createdAt), 
         updatedAt = date(updatedAt))

if(remove_testing){
  studies_clean <- studies_clean %>% filter(!projectId %in% c('testing-research', 'testing-teaching'))
}

```

```{r include = FALSE}
studyperm_clean <- studyperm %>%
  select(id, adminAccess, adminUsers, createdAt, readonlyAccess, readonlyUsers, readwriteAccess, readwriteUsers,
         recordType, uid, writeonlyAccess, writeonlyUsers) %>%
    mutate(createdAt = date(createdAt),
         adminAccess = clean_list(adminAccess, sep = ':', rep = 'XXX', newsep = ', '),
         adminUsers = clean_list(adminUsers, sep = ':', rep = 'XXX', newsep = ', '),
         readonlyAccess = clean_list(readonlyAccess, sep = ':', rep = 'XXX', newsep = ', '),
         readonlyUsers = clean_list(readonlyUsers, sep = ':', rep = 'XXX', newsep = ', '),
         readwriteAccess = clean_list(readwriteAccess, sep = ':', rep = 'XXX', newsep = ', '),
         readwriteUsers = clean_list(readwriteUsers, sep = ':', rep = 'XXX', newsep = ', '),
         writeonlyAccess = clean_list(writeonlyAccess, sep = ':', rep = 'XXX', newsep = ', '),
         writeonlyUsers = clean_list(writeonlyUsers, sep = ':', rep = 'XXX', newsep = ', '))

study_only_perm <- studyperm_clean %>%
  filter(recordType == 'study') %>%
  left_join(user_mapping, by = c('adminUsers' = 'uid')) %>%
  rename(adminUsernames = username) %>%
  left_join(user_mapping, by = c('readonlyUsers' = 'uid')) %>%
  rename(readonlyUsernames = username) %>%
  left_join(user_mapping, by = c('readwriteUsers' = 'uid')) %>% 
  rename(readwriteUsernames = username) %>%
  left_join(user_mapping, by = c('writeonlyUsers' = 'uid')) %>%
  rename(writeonlyUsernames = username) %>%
  select(id, createdAt, adminUsernames, readonlyUsernames, readwriteUsernames, writeonlyUsernames) %>%
  mutate(id = gsub('Study:', '', id))

users_only_perm <- studyperm_clean %>%
  filter(recordType == 'user') %>% 
  left_join(users_clean %>% select(uid, username), by = c('uid' = 'uid')) %>%
  select(id, username, adminAccess, readonlyAccess, readwriteAccess, writeonlyAccess)

```

```{r echo = FALSE}
# Service Workbench Metrics report

print(paste0('generated ', date()))

```


# User Summary

### How many users are in Service Workbench?

Active users have been approved by an administrator and assigned to a project.

Inactive users are pending approval by an administrator. 



```{r echo = FALSE}
# users by status
# TABLE
# users_clean %>%
#   group_by(status) %>%
#   mutate(status = gsub('pending', 'inactive', status)) %>%
#   summarise(n = n_distinct(uid)) %>% 
#   kable(caption = 'Number of users by status',
#         col.names = c("User Status", "Number of Users"), 
#         align = 'l')

# PIE CHART
users_clean %>%
  group_by(status) %>%
  mutate(status = gsub('pending', 'inactive', status)) %>%
  summarise(n = n_distinct(uid)) %>% 
  ggplot(aes(x="", y=n, fill=status)) +
    geom_col() +
    geom_text(aes(label = paste0(n, ' users')),
            position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    theme_void() +
    labs(fill = "User Status",
         title = "Number of users by status",
         subtitle = paste0(n_distinct(users_clean$uid), ' total users'))
  

```

\newpage

### What projects are our users working on?

Service Workbench organizes users into *projects*. These allow us to categorize users, organize billing for computational resources and storage, and see at a high level who is using the platform.

In this report, projects are further organized into *project groups* to demonstrate the main user groups in AIM AHEAD SWB. The Project ID references the more granluar project designation.

Users may be assigned to multiple projects. Because AIM-AHEAD non-fisma SWB was our first test of the SWB infrastructure in AIM-AHEAD, we have many users assigned to the *testing-teaching* and *testing-research* projects. 
```{r include = FALSE}
# users_clean %>%
#   select(projectId, username) %>%
#   separate_rows(projectId, sep = ', ') %>%
#   left_join(project_groupings) %>%
#   mutate(Project_Group = ifelse(is.na(Project_Group), 'Unassigned', Project_Group)) %>%
#   group_by(Project_Group) %>%
#   summarise(n = n()) %>%
#   ggplot(aes(x="", y=n, fill=Project_Group)) +
#     geom_col() +
#     geom_text(aes(label = paste0(n, ' users')),
#             position = position_stack(vjust = 0.5)) +
#     coord_polar(theta = "y") +
#     theme_void() +
#     labs(fill = "User Projects",
#          title = "Number of users by project",
#          subtitle = paste0(n_distinct(users_clean$uid), ' total users'))

```

```{r echo = FALSE}
# t <- users_clean %>%
#   select(projectId, username) %>%
#   separate_rows(projectId, sep = ', ') %>%
#   left_join(project_groupings) %>%
#   mutate(projectId = ifelse(projectId == '', 'Unassigned', projectId)) %>%
#   mutate(Project_Group = ifelse(is.na(Project_Group), 'Unassigned', Project_Group)) %>%
#   group_by(projectId) %>%
#   mutate(n_users = n_distinct(username)) %>% 
#   arrange(desc(n_users)) %>%
#   select(-username) %>%
#   unique()
# t$projectId <- factor(t$projectId, levels = unique(t$projectId))
# t$Project_Group <- factor(t$Project_Group, levels = unique(t$Project_Group))
# t %>%
#   ggplot(aes(x = projectId, y = n_users, fill = Project_Group)) +
#     geom_bar(stat = 'identity') +
#     geom_text(aes(label=n_users), vjust = -.5) + 
#     theme_bw() +
#     theme(axis.text.x = element_text(angle = 45,  hjust=1)) +
#     labs(x = 'Project ID', y = 'Number of Users', fill = 'Project Group',
#          title = 'Number of unique users per project', 
#          subtitle = "Note that users may be assigned to multiple projects") +
#     ylim(c(0, max(t$n_users) + 10))
  
```

```{r echo = FALSE}
t <- users_clean %>%
  select(projectId, username) %>%
  separate_rows(projectId, sep = ', ') %>%
  left_join(project_groupings) %>%
  mutate(Project_Group = ifelse(is.na(Project_Group), 'Unassigned', Project_Group)) %>%
  group_by(Project_Group) %>%
  summarise(n_users = n_distinct(username)) %>% 
  arrange(desc(n_users)) %>%
  unique()
t$Project_Group <- factor(t$Project_Group, levels = unique(t$Project_Group))
t %>%
  ggplot(aes(x = Project_Group, y = n_users, fill = Project_Group)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label=n_users), vjust = -.5) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45,  hjust=1), 
          legend.position = 'none') + 
    labs(x = 'Project Group', y = 'Number of Users',
         title = 'Number of unique users per project group', 
         subtitle = "Note that users may be assigned to multiple projects") +
    ylim(c(0, max(t$n_users) + 10))
  
```


### Detailed breakdown of Project Groups

For more information on what projects make up the Project Groups visualized in the above chart, see the table below:

```{r echo = FALSE}
project_groupings %>% 
  arrange(Project_Group) %>%
  kable(row.names = FALSE)

```


\newpage

# Workspace Summary

### What types of workspaces are being used?

Workspaces are the foundation of a user's interaction with Service Workbench. 

This workspace breakdown demonstrates which workspaces are most used. 

Sagemaker tends to be the most popular workspace for researchers as it provides a simple, Jupyter Notebook interface where users can code in R or python.




```{r echo = FALSE}
workspaces_clean %>%
  mutate(type = gsub(x = envTypeConfigId, '-.*$|_.*$', ''), 
         type = str_to_upper(type), 
                  type = ifelse(grepl('RSTUDIO', type), 'RSTUDIO', 
                             ifelse(grepl('SAGEMAKER', type), 'SAGEMAKER', 
                                          ifelse(grepl('WINDOWS', type), 'WINDOWS', 
                                                       ifelse(grepl('LINUX', type), 'LINUX', 'OTHER'))))) %>%
  group_by(type) %>%
  summarise(n_workspaces = n_distinct(id)) %>%
  ggplot(aes(x="", y=n_workspaces, fill=type)) +
    geom_col() +
    geom_text(aes(label = n_workspaces),
            position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    theme_void() +
    labs(fill = "Workspace Type",
         title = "Number of workspaces created",
         subtitle = paste0(n_distinct(workspaces_clean$id), ' total workspaces')) 

```

\newpage

# Which users are using which workspaces?

The following chart shows the breakdown of workspaces per project. 

The large amount of testing workspaces demonstrates the constant need for QA as we develop new features and maintain the infrastructure services. 

We can see that AIM AHEAD Fellowship users tend to prefer Sagemaker workspaces while users using SWB for teaching tend to prefer RStudio workspaces.


```{r echo = FALSE}

w <- workspaces_clean %>%
  mutate(type = gsub(x = envTypeConfigId, '-.*$|_.*$', ''), 
         type = str_to_upper(type),
         type = ifelse(grepl('RSTUDIO', type), 'RSTUDIO', 
                             ifelse(grepl('SAGEMAKER', type), 'SAGEMAKER', 
                                          ifelse(grepl('WINDOWS', type), 'WINDOWS', 
                                                       ifelse(grepl('LINUX', type), 'LINUX', 'OTHER'))))) %>%
  group_by(type, Project_Group) %>%
  summarise(n_workspaces = n_distinct(id)) %>%
  arrange(desc(n_workspaces))

ww <- w %>% group_by(Project_Group) %>% summarise(n = sum(n_workspaces)) %>% arrange(desc(n))

w$Project_Group <- factor(w$Project_Group, levels = ww$Project_Group)

w %>%
  ggplot(aes(x = Project_Group, y=n_workspaces, fill=type)) +
    geom_col() +
    geom_text(aes(label = n_workspaces), position = position_stack(vjust = 0.5)) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = 'Project Group', y = 'Number of Users', fill = "Workspace Type",
          title = 'Number of workspaces per project', 
          subtitle = paste0(n_distinct(workspaces_clean$id), ' total workspaces')) 



```

\newpage

# Computational Resource Summary

The following summaries represent the computational resources provided & used by Service Workbench users. 

This is a rough estimate based on workspace types. A more accurate representation based on instance types is in development for 2024. 

Computational resources provided per group:
```{r echo = FALSE}
cr <- workspaces_clean %>%
  group_by(Project_Group) %>% 
  summarise(n_CPU = sum(CPU, na.rm = TRUE),n_GPU = sum(GPU, na.rm = TRUE),n_RAM = sum(RAM, na.rm = TRUE)) 

cr %>% 
  kable(col.names = c("Project Group", "Number of CPUs", "Number of GPUs", "Amount of RAM (GiB)")) 


paste0("In total, AIM AHEAD non-FISMA SWB made ", 
       sum(cr$n_CPU, na.rm = TRUE), " CPUs, ", 
       sum(cr$n_GPU, na.rm = TRUE), " GPUs, and ", 
       sum(cr$n_RAM, na.rm = TRUE), " GiB RAM",
       " available to users.")

```







