---
title: "SWB Metrics Report"
subtitle: "GIC"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---



```{r include = FALSE}
#Download all entries from the following DynamoDB tables: 
#- aim-prod-va-sw-Users
#- aim-prod-va-sw-EnvironmentsSc
#- Studies
#- StudyPermissions

users <- read.csv('../db-exports/gic-users.csv')
workspaces <- read.csv('../db-exports/gic-environments.csv')
studies <- read.csv('../db-exports/gic-studies.csv')
studyperm <- read.csv('../db-exports/gic-studypermissions.csv')

library(tidyverse)
library(knitr)

clean_list <- function(vec, sep, rep, newsep){
  vec = gsub(pattern = sep, replacement = rep, x = vec)
  vec = gsub(pattern = '[[:punct:]]', replacement = '', x = vec)
  vec = gsub(pattern = paste0('S', rep), replacement = newsep, x = vec)
  vec = gsub(pattern = paste0('^', newsep), replacement = '', x = vec)
  return(vec)
}

# keep testing to show behind the scenes work
remove_testing = FALSE

```


```{r include = FALSE}

users_clean <- users %>% 
  select(uid, firstName, lastName, email, username, applyReason, 
         projectId, status,  userRole,
         createdAt, acceptedTerms, identityProviderName, isAdmin) %>%
  mutate(createdAt = date(createdAt), 
         acceptedTerms = ifelse(acceptedTerms == '', NA, acceptedTerms),
         acceptedTerms = date(acceptedTerms),
         projectId = clean_list(projectId, sep = ':', rep = 'XXX', newsep = ', ')) 

user_mapping <- users_clean %>% 
  select(uid, username) %>%
  mutate(uid = clean_list(uid, '', '', ''))

```

```{r include = FALSE}

workspaces_clean <- workspaces %>% 
  select(id, name, createdAt, createdBy, description, outputs,
         envTypeConfigId, indexId,  projectId, status, studyIds, updatedAt, instanceType) %>%
  mutate(createdAt = date(createdAt),
         updatedAt = date(updatedAt), 
         projectId = gsub(pattern = '[[:punct:]]', replacement = '', x = projectId),
         studyIds = clean_list(studyIds, sep = ':', rep = 'XXX', newsep = ', '),
         envTypeConfigId = tolower(envTypeConfigId), 
         envTypeConfigId = clean_list(envTypeConfigId, sep = '', rep = '', newsep = '')) %>%
  left_join(users_clean %>% select(uid, username), by = c('createdBy' = 'uid')) 

if(remove_testing){
  workspaces_clean <- workspaces_clean %>% filter(Project_Group != "Testing")
}


```

```{r include = FALSE}
studies_clean <- studies %>%
  select(id, accessType, bucket, category, createdAt, description, folder, name, projectId, updatedAt) %>%
  mutate(createdAt = date(createdAt), 
         updatedAt = date(updatedAt))

if(remove_testing){
  studies_clean <- studies_clean %>% filter(!projectId %in% c('testing-research', 'testing-teaching'))
}

```

```{r include = FALSE}
studyperm_clean <- studyperm %>%
  select(id, adminAccess, adminUsers, createdAt, readonlyAccess, readonlyUsers, readwriteAccess, readwriteUsers,
         recordType, uid, writeonlyAccess, writeonlyUsers) %>%
    mutate(createdAt = date(createdAt),
         adminAccess = clean_list(adminAccess, sep = ':', rep = 'XXX', newsep = ', '),
         adminUsers = clean_list(adminUsers, sep = ':', rep = 'XXX', newsep = ', '),
         readonlyAccess = clean_list(readonlyAccess, sep = ':', rep = 'XXX', newsep = ', '),
         readonlyUsers = clean_list(readonlyUsers, sep = ':', rep = 'XXX', newsep = ', '),
         readwriteAccess = clean_list(readwriteAccess, sep = ':', rep = 'XXX', newsep = ', '),
         readwriteUsers = clean_list(readwriteUsers, sep = ':', rep = 'XXX', newsep = ', '),
         writeonlyAccess = clean_list(writeonlyAccess, sep = ':', rep = 'XXX', newsep = ', '),
         writeonlyUsers = clean_list(writeonlyUsers, sep = ':', rep = 'XXX', newsep = ', '))

study_only_perm <- studyperm_clean %>%
  filter(recordType == 'study') %>%
  left_join(user_mapping, by = c('adminUsers' = 'uid')) %>%
  rename(adminUsernames = username) %>%
  left_join(user_mapping, by = c('readonlyUsers' = 'uid')) %>%
  rename(readonlyUsernames = username) %>%
  left_join(user_mapping, by = c('readwriteUsers' = 'uid')) %>% 
  rename(readwriteUsernames = username) %>%
  left_join(user_mapping, by = c('writeonlyUsers' = 'uid')) %>%
  rename(writeonlyUsernames = username) %>%
  select(id, createdAt, adminUsernames, readonlyUsernames, readwriteUsernames, writeonlyUsernames) %>%
  mutate(id = gsub('Study:', '', id))

users_only_perm <- studyperm_clean %>%
  filter(recordType == 'user') %>% 
  left_join(users_clean %>% select(uid, username), by = c('uid' = 'uid')) %>%
  select(id, username, adminAccess, readonlyAccess, readwriteAccess, writeonlyAccess)

```

```{r echo = FALSE}
# Service Workbench Metrics report

print(paste0('generated ', date()))

```


# User Summary

### How many users are in Service Workbench?

Active users have been approved by an administrator and assigned to a project.

Inactive users may be pending project approval. 



```{r echo = FALSE}
# users by status
# TABLE
# users_clean %>%
#   group_by(status) %>%
#   mutate(status = gsub('pending', 'inactive', status)) %>%
#   summarise(n = n_distinct(uid)) %>% 
#   kable(caption = 'Number of users by status',
#         col.names = c("User Status", "Number of Users"), 
#         align = 'l')

# PIE CHART
users_clean %>%
  group_by(status) %>%
  mutate(status = gsub('pending|inactive', 'Pending', status),
         status = gsub('active', 'Active', status)) %>%
  summarise(n = n_distinct(uid)) %>% 
  ggplot(aes(x="", y=n, fill=status)) +
    geom_col() +
    geom_text(aes(label = paste0(n, ' users')),
            position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    theme_void() +
    labs(fill = "User Status",
         title = "Number of users by status",
         subtitle = paste0(n_distinct(users_clean$uid), ' total users'))
  
# print users who are pending
print('pending users:')
users_clean %>% 
  filter(status %in% c('pending', 'inactive')) %>% 
  select(firstName, lastName, email, projectId, userRole, createdAt) %>%
  arrange(desc(createdAt)) %>%
  kable()

```

\newpage

### What institutions are our users part of?

In GIC SWB, projects reflect institutions within GIC.

```{r echo = FALSE}
# Leap-dev users are being counted as BCH users as this was a BCH led project. 


proj_mapping <- data.frame('pid' = c('bchresearch', 'GICWASHU', 'GICCCHMC', 'GICCHOP', 'GICUTHSC', 'LEAPDEV'),
                           'site' = c('BCH', 'WASHU', 'CCHMC', 'CHOP', 'UTHSC', 'BCH'))

t <- users_clean %>%
  filter(userRole == 'researcher') %>% # remove admins and unapproved users
  select(projectId, username) %>%
  separate_rows(projectId, sep = ', ') %>%
  left_join(proj_mapping, by = c('projectId' = 'pid')) %>%
  group_by(site) %>%
  summarise(n_users = n_distinct(username)) %>% 
  arrange(desc(n_users)) %>%
  unique()
t$site <- factor(t$site, levels = unique(t$site))
t %>%
  ggplot(aes(x = site, y = n_users, fill = site)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label=n_users), vjust = -.5) + 
    theme_bw() +
    theme(legend.position = 'none') + 
    labs(x = 'Institution', y = 'Number of Users',
         title = 'Number of unique users per institution') +
    ylim(c(0, max(t$n_users) + 10)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30))
  
```




### User growth over time
```{r echo = FALSE}
user_growth <- users_clean %>% 
  select(username, createdAt) %>%
  mutate(month_created = format(createdAt, "%b %Y"))

min_date <- min(user_growth$createdAt)
max_date <- max(user_growth$createdAt)
dates <- seq.Date(min_date, max_date, by = 'day') %>% format('%b %Y') %>% unique() %>%
  factor(seq.Date(min_date, max_date, by = 'day') %>% format('%b %Y') %>% unique())
add <- data.frame(month_created = dates[!dates %in%user_growth$month_created],
                  n = 0)

user_growth_summary <- user_growth %>%
  group_by(month_created) %>%
  summarise(n = n())%>%
  rbind(add) %>%
  arrange(factor(month_created, levels = dates))

user_growth_summary$month_created <- factor(user_growth_summary$month_created, levels = dates)


ggplot(user_growth_summary) +
  geom_bar(aes(x = month_created, y = n), stat = 'identity') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.95), 
          legend.position = 'none') +
  labs(x = 'Month', y = 'Number of Users Added',
         title = 'Number of users added per Month')

ggplot(user_growth_summary) +
  geom_bar(aes(x = month_created, y = cumsum(n)), stat = 'identity') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.95), 
          legend.position = 'none') +
  labs(x = 'Month', y = 'Number of Users',
         title = 'Number of total users over time')


```


\newpage

# Workspace Summary

### What types of workspaces are being used?

Workspaces are the foundation of a user's interaction with Service Workbench. 

- **SageMaker** workspaces provide users with a Jupyter Notebook interface where they have access to the following:
  - User-friendly file explorer
  - R coding in Jupyter Notebooks
  - python coding in Jupyter Notebooks
  - Terminal access
  
- **Rstudio** workspaces provide users who code in R an Rstudio interface to easily write analyses. This workspace type also provides users with access to:
  - User-friendly file explorer
  - R coding in .R or .Rmd files
  - Terminal access
  
- **Linux** workspaces provide users access to ssh into a linux machine. Users can also use linux workspaces to upload large files to S3 storage. 

- **Windows** workspaces provide users acess to a Windows virtual desktop. These workspaces are frequently used by users accessing OCHIN data, as they provide access to the SQL Server Management Studio app that allows users to investigate and query the patient level data. 

This workspace breakdown demonstrates which workspaces are most used. 

Sagemaker tends to be the most popular workspace for researchers as it provides a simple, Jupyter Notebook interface where users can code in R or python.




```{r echo = FALSE}
workspaces_clean %>%
  mutate(type = gsub(x = envTypeConfigId, '-.*$|_.*$', ''), 
         type = str_to_upper(type), 
                  type = ifelse(grepl('RSTUDIO', type), 'RSTUDIO', 
                             ifelse(grepl('SAGEMAKER|RAPIDS|TEST', type), 'SAGEMAKER', 
                                          ifelse(grepl('WINDOWS', type), 'WINDOWS', 
                                                       ifelse(grepl('LINUX', type), 'LINUX', 'OTHER'))))) %>%
  group_by(type) %>%
  summarise(n_workspaces = n_distinct(id)) %>%
  ggplot(aes(x="", y=n_workspaces, fill=type)) +
    geom_col() +
    geom_text(aes(label = n_workspaces),
            position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    theme_void() +
    labs(fill = "Workspace Type",
         title = "Number of workspaces by type",
         subtitle = paste0(n_distinct(workspaces_clean$id), ' total workspaces')) 
 
```








